Processing genome scale tabular data with wormtable

### 2013-11-19 docs/paper/paper.tex
Addressed comments from editor and reviewers.

The classical approach to solving problems of this type 
is to use a relational database, which provide sophisticated 
data management techniques. However, relational databases are unsuitable 
for storing static datasets as they are complex to 
use and incur many unnecessary overheads.
Wormtable provides the most important features of database
technologies (packed binary storage of values; random access 
to rows; general purpose indexing) without additional complexities 
and overheads. Wormtable is also more flexible than 
relational databases, supporting, for example, columns 
containing a variable number of integers. Finally, wormtable 
is far more adaptable than a relational database.
All widely used database systems are complex 
and adding the necessary features would be 
very difficult. Adding new features to wormtable, on the 
other hand, is straightforward because it is far simpler and 
does not need to be compatible with the relational model
and decades worth of existing software.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Authors contributions}
JK conceived the project, developed wormtable source code and associated
scripts and drafted the manuscript. RN wrote the tutorial and contributed to
the example scripts and the manuscript. DH wrote the example scripts and vcf-utils,
and contributed to vcf2wt, the tutorial and the manuscript.
All authors read and approved the final manuscript.

>>>>

The classical approach to solving problems of this type 
is to use a relational database, which provide sophisticated 
data management techniques. However, relational databases are unsuitable 
for storing static datasets as they are complex to 
use and incur many unnecessary overheads.
Wormtable provides the most important features of database
technologies (packed binary storage of values; random access 
to rows; general purpose indexing) without additional complexities 
and overheads. Wormtable's data model is also less rigid
than 
relational databases, supporting, for example, columns 
containing a variable number of integers. Finally, wormtable 
is far more adaptable than a relational database.
All widely used database systems are complex 
and adding required features (e.g.\ compression) would be 
very difficult. Adding new features to wormtable, on the 
other hand, is straightforward because it is far simpler and 
does not need to be compatible with the relational model
and decades worth of existing software.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Competing interests}
The authors declare that they have no competing interests.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section*{Authors contributions}
JK conceived the project, developed wormtable source code and associated
scripts and drafted the manuscript. RN wrote the tutorial and contributed to
the example scripts and the manuscript. DH wrote the example scripts and vcf-utils,
and contributed to \texttt{vcf2wt}, the tutorial and the manuscript.
All authors read and approved the final manuscript.
